Exploring probability distributions for bivariate temporal granularities

### 2019-11-24 main.Rmd
all notations change in the aperiodic definition

**Definition: Aperiodic granularity** An aperiodic granularity A(G, BG) that relates linear granularities G and bottom granularity BG, with G being periodical with respect to BG, is given by

$A_{BG, G}(z) = BG(z - \sum_{j=0}^{k-1}\vert S_{j \mod R} \vert)$ for $z \in S_k$,

where,
z denotes the index set,  
R denotes the number of granules of G in each period of BG,  
$S_j$ are the sets of indices of BG describing $G_j$ such that $G(j) = \bigcup_{i \in S_j}BG(i)$,  
$\vert S_j \vert$ is the cardinality of set $S_j$,

Since G is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:

- $BG \trianglelefteq G,$
- $G(j) = \bigcup_{i \in S_j \mod R}BG(P*\lfloor j/R \rfloor + i)$,  where P denotes the period of the granularity BG.

**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the diagram \autoref{fig:aperiodic-gran}. Each granule is represented by a box. Similar to Figure \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity. 
Two linear granularities G1 and G2 are considered, where $G1 = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $G2 = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$. This implies that G1 is made up of shrinking every 1st granule of Group_3(BG) by 1 and G2 is made up of shrinking every 1st granule of Group_3(BG) by 2. Number of granules of G1 and G2 in each period of BG is 2, but the number of granules of BG in each of those granules are different. A(G, G1) and A(G, G2) are repetitive categorisation of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of G.

@@

To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar.

@@

<!-- , and indeed many parts of the world. Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning. The behavioral patterns in households vary substantially, for example, some families use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.  -->

>>>>

**Definition: Aperiodic granularity** An aperiodic granularity A(AG, BG) that relates linear granularities AG and bottom granularity BG, with AG being periodical with respect to BG, is given by

$A_{BG, AG}(z) = BG(z - \sum_{w=0}^{k-1}\vert T_{w \mod R'} \vert)$ for $z \in T_k$,

where,  
z denotes the index set (index of BG),
R' denotes the number of granules of AG in each period of BG,  
$T_w$ are the sets of indices of BG describing $AG_w$ such that $AG(w) = \bigcup_{z \in T_w}BG(z)$,  
$\vert T_w \vert$ is the cardinality of set $T_w$,

Since AG is periodical with respect to BG, from definition of "Groups periodically into" the following holds true:

- $BG \trianglelefteq AG,$
- $AG(w) = \bigcup_{z \in T_w \mod R}BG(P'*\lfloor w/R' \rfloor + z)$,  where P' denotes the period of the granularity BG.

**Discussion**: Example showing aperiodic granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in \autoref{fig:aperiodic-gran}. Each granule is represented by a box. Similar to \autoref{fig:circular-gran}, the granules that overlap share elements from the underlying time domain and the first slot shows the index set and the bottom granularity. 
Two linear granularities AG and AH are considered, where $AG = Alter^{5}_{(1,-1)}(BG, Group_3(BG))$ and $AH = Alter^{12}_{(1,-2)}(BG, Group_7(BG))$. This implies that AG is made up of shrinking every 1st granule of Group_3(BG) by 1 and AH is made up of shrinking every 1st granule of Group_3(BG) by 2. Number of granules of AG and AH in each period of BG is 2, but the number of granules of BG in each of those granules are different. A(BG, AG) and A(BG, AH) are repetitive categorisation of time, similar to circular granularities, except that the number of granules of BG is not necessarily the same across different granules of AG or AH. For AG, R' = 2, P'= 5, $T_0$ = {0,1} and $T_1$ = {2, 3, 4}, suppose we want to compute A(BG,AG)(8), then since $8 \in T_{3}$

\begin{equation} \label{eq4}
\begin{split}
A(BG,AG)(8) & = BG(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) \\
  & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
  & = BG(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
  & = BG(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
  & = BG(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
  & = BG(8 - 2.2 - 3) \\
  & = BG(1) \\
\end{split}
\end{equation}

@@

The linear granularities G, H, G1 and G2 considered in the definitions of both circular and aperiodic granularities are periodic with respect to the bottom granularity. The only difference is G, H have regular mapping, whereas, G1 and G2 have irregular mapping with respect to BG.

<!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->

@@

<!-- , and indeed many parts of the world. Households are distributed geographically and have different demographic properties such as the existence of solar panels, central heating or air conditioning. The behavioral patterns in households vary substantially, for example, some famailies use a dryer for their clothes while others hang them on a line, and some households might consist of night owls, while others are morning larks.  -->
